---
title: "Regression Segment improved"
author: '220455224'
date: "`r Sys.Date()`"
output: html_document
---

The data set we are using to showcase regression techniques is 'Global Food Wastage' from Kaggle. 
Link: https://www.kaggle.com/datasets/atharvasoundankar/global-food-wastage-dataset-2018-2024
After downloading the data set, set working directory to where the file is located.
```{r}
#change path to where the csv file is located
setwd("C:/Machine Learning")
```
Now, let us tackle this question, "Can we predict economic loss due to food waste based on total waste, average waste per capita, population and household waste?" using regression techniques.

Install the packages required
```{r}
install.packages("tidyverse")
install.packages("caret")
install.packages("glmnet")
install.packages("randomForest")
```
Load the packages
```{r}
library(tidyverse)
library(caret)
library(glmnet)
library(randomForest)
```
Read the csv file into a dataframe
```{r}
Fooddf <- read.csv("global_food_wastage_dataset.csv")
```
Check structure of dataframe
```{r}
str(Fooddf)
```
Rename columns
```{r}
colnames(Fooddf) <- c("Country", "Year", "Food_Category", "Total_Waste_Tons", 
                       "Economic_Loss_Million", "Avg_Waste_per_Capita_Kg", 
                       "Population_Million", "Household_Waste_Percent")

print(colnames(Fooddf))
```
Remove rows with NA values
```{r}
Fooddf <- na.omit(Fooddf)
head(Fooddf)
```
Checking the correlation
```{r}
cor(Fooddf[, c("Total_Waste_Tons", "Avg_Waste_per_Capita_Kg", "Population_Million", "Household_Waste_Percent")])
```
Scale the data
```{r}
Fooddf_scaled <- Fooddf
Fooddf_scaled[, c("Total_Waste_Tons", "Avg_Waste_per_Capita_Kg", "Population_Million", "Household_Waste_Percent")] <- scale(Fooddf[, c("Total_Waste_Tons", "Avg_Waste_per_Capita_Kg", "Population_Million", "Household_Waste_Percent")])
head(Fooddf_scaled)
```
Select features that we want to use and split data into training and testing sets
```{r}
set.seed(123) 
trainIndex <- createDataPartition(Fooddf$Economic_Loss_Million, p = 0.8, list = FALSE)
#Split
train <- Fooddf_scaled[trainIndex, ]
test <- Fooddf_scaled[-trainIndex, ]
#Check the rows in dataframe
head(train)
```
Multiple Linear regression
```{r}
lr_model <- lm(Economic_Loss_Million ~ Total_Waste_Tons + Avg_Waste_per_Capita_Kg + 
               Population_Million + Household_Waste_Percent, data = train)

summary(lr_model)
```
Plot a residual plot
```{r}
plot(lr_model$residuals)
#Add a horizontal line at 0 to see the residuals' distribution
abline(h = 0, col = "red")
```
Random Forest
```{r}
rf_model <- randomForest(Economic_Loss_Million ~ Total_Waste_Tons +         Avg_Waste_per_Capita_Kg + Population_Million + Household_Waste_Percent, data = train, ntree = 500)
#See model's performance
print(rf_model) 
```
Models evaluation
```{r}
#Each model predictions
lr_pred <- predict(lr_model, test)
rf_pred <- predict(rf_model, test)
#Compute RMSE for each model
lr_rmse <- sqrt(mean((lr_pred - test$Economic_Loss_Million)^2))
rf_rmse <- sqrt(mean((rf_pred - test$Economic_Loss_Million)^2))
#Print the results
cat("Linear Regression RMSE:", lr_rmse, "\n")
cat("Random Forest RMSE:", rf_rmse, "\n")
```
Plot predicted vs actual economic loss values
```{r}
#Retrieve predicted values from linear regression model
predictions <- predict(lr_model, newdata = test)

#Plot predicted vs actual economic loss
ggplot(data = NULL, aes(x = test$Economic_Loss_Million, y = predictions)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", color = "red") +
  labs(x = "Actual Economic Loss ($ Million)", y = "Predicted Economic Loss ($ Million)",
       title = "Predicted vs Actual Economic Loss") +
  theme_minimal()
```
To determine whether the three other variables were strongly contributing to or being absorbed into total waste. 
```{r}
#Scale
Fooddf_scaled2 <- Fooddf
Fooddf_scaled2[, c("Avg_Waste_per_Capita_Kg", "Population_Million", "Household_Waste_Percent")] <- scale(Fooddf[, c("Avg_Waste_per_Capita_Kg", "Population_Million", "Household_Waste_Percent")])
#Split the data
set.seed(123) 
trainIndex2 <- createDataPartition(Fooddf$Total_Waste_Tons, p = 0.8, list = FALSE)
#Split
train2 <- Fooddf_scaled2[trainIndex2, ]
test2 <- Fooddf_scaled2[-trainIndex2, ]
#Check the rows in dataframe
head(train2)
#Train the model with total waste as the target variable
lr_model2 <- lm(Total_Waste_Tons ~ Avg_Waste_per_Capita_Kg + 
                 Population_Million + Household_Waste_Percent, data = train2)

summary(lr_model2)
```