---
title: "220455224 Classification segment"
author: '220455224'
date: "`r Sys.Date()`"
output: html_document
---

The data set we are using to showcase classification techniques is 'Resale Flat Prices' from data.gov.sg. 
Link: https://data.gov.sg/datasets?page=3&query=&formats=CSV&resultId=d_ebc5ab87086db484f88045b47411ebc5
After downloading the data set, set working directory to where the file is located.
```{r}
#change path to where the csv file is located
setwd("C:/Machine Learning")
```
Install the packages required
```{r}
install.packages("dplyr")
install.packages("rpart")
install.packages("rpart.plot")
install.packages("caTools")
install.packages("nnet")
install.packages("caret")
```
Load the packages
```{r}
library(dplyr)
library(rpart)
library(rpart.plot)
library(caTools)
library(nnet)
library(caret)
```
Research Question: Can we predict the flat model based on features like floor area, resale price, lease commence year and flat type?

Read the csv file into a dataframe
```{r}
Rdf <- read.csv("ResaleFlatPricesBasedonApprovalDate19901999.csv")
```
Extract the year we want to use
```{r}
#Extract year from the 'month' column
Rdf$year <- as.numeric(substr(Rdf$month, 1, 4))

#Filter for years 1992 to 1993
Rdf_filtered <- Rdf %>%
  filter(year >= 1992 & year <= 1993)
```
Check for any missing values
```{r}
#Check total missing values
sum(is.na(Rdf_filtered))
#Check per column
colSums(is.na(Rdf_filtered))
```
Convert relevant categorical variables into factors
```{r}
Rdf_filtered$flat_model <- as.factor(Rdf_filtered$flat_model)
Rdf_filtered$flat_type <- as.factor(Rdf_filtered$flat_type)
head(Rdf_filtered)
```
Select variables that might influence the model
```{r}
Rdf_filtered <- Rdf_filtered %>%
  select(flat_model, floor_area_sqm, resale_price, flat_type, lease_commence_date)
head(Rdf_filtered)
```
Split data sets into training and testing sets
```{r}
set.seed(123) 
split <- sample.split(Rdf_filtered$flat_model, SplitRatio = 0.7)
#Split
train_data <- subset(Rdf_filtered, split == TRUE)
test_data <- subset(Rdf_filtered, split == FALSE)
#Check for missing values
colSums(is.na(train_data))
```
Train Decision Tree model
```{r}
tree_model <- rpart(flat_model ~ ., data = train_data, method = "class")
```
Plot the decision tree
```{r}
rpart.plot(tree_model)
```
Pruning the tree and save the decision tree into a png for better viewing
```{r}
#Pruning tree
tree_model <- rpart(flat_model ~ ., data = train_data, method = "class", cp = 0.02)  
#Saving decision tree in png 
png("decision_tree.png", width = 2000, height = 1200)
rpart.plot(tree_model, type = 3, tweak = 1.2, extra = 104)  
dev.off()
#Plot decision tree in R
rpart.plot(tree_model, type = 3, tweak = 1.2, extra = 104)  
```
Multinomial Logistic Regression
```{r}
logistic_model <- multinom(flat_model ~ ., data = train_data)
#Show statistics
summary(logistic_model)
```
Find out the null deviance in the multinomial logistic regression model
```{r}
null_model <- multinom(flat_model ~ 1, data = train_data)
summary(null_model)
```
Evaluate performance for Decision Tree using Confustion Matrix
```{r}
tree_pred <- predict(tree_model, test_data, type = "class")
confusionMatrix(tree_pred, test_data$flat_model)
```

Confusion matrix for Multinomial Logistic Regression
```{r}
logistic_pred <- predict(logistic_model, test_data, type = "class")
confusionMatrix(logistic_pred, test_data$flat_model)
```