---
title: " 220455224 Unsupervised learning segment"
author: '220455224'
date: "`r Sys.Date()`"
output: html_document
---

The data set we are using to showcase unsupervised learning techniques is 'Resale Flat Prices' from data.gov.sg. 
Link: https://data.gov.sg/datasets?page=3&query=&formats=CSV&resultId=d_ebc5ab87086db484f88045b47411ebc5
After downloading the data set, set working directory to where the file is located.
```{r}
#change path to where the csv file is located
setwd("C:/Machine Learning")
```
Install the packages required
```{r}
install.packages("dplyr")
install.packages("cluster")
install.packages("factoextra")
install.packages("ggplot2")
```
Load the packages
```{r}
library(dplyr)
library(factoextra)
library(ggplot2)
library(cluster)
```
Read the csv file into a dataframe
```{r}
df <- read.csv("ResaleFlatPricesBasedonApprovalDate19901999.csv")
```
Research question: Can we cluster the houses by flat types? How does flat type affect variables?

Filter for years 1990 to 1991.
```{r}
#Extract year from the 'month' column
df$year <- as.numeric(substr(df$month, 1, 4))
#Filter for years 1990 to 1991
df_filtered <- df %>%
  filter(year >= 1990 & year <= 1991)
```
Check number of flat types
```{r}
unique(df_filtered$flat_type)
```
Convert flat type to numeric
```{r}
df_filtered$flat_type_numeric <- as.numeric(factor(df_filtered$flat_type, levels = unique(df_filtered$flat_type)))
```
Select variables that will be used and remove missing values if present
```{r}
clustdf <- df_filtered %>%
  select(flat_type_numeric, floor_area_sqm, resale_price) %>%
  na.omit()
```
Take a random sample of 10000
```{r}
set.seed(123) 
sampledf<- clustdf %>% sample_n(min(n(), 10000)) 
```
Normalise the data
```{r}
scaledsampledf <- as.data.frame(scale(sampledf))
```
Find out optimal number of clusters
```{r}
fviz_nbclust(scaledsampledf, kmeans, method = "wss")
```
Using k-means clustering
```{r}
set.seed(123)
k <- 3
km_result <- kmeans(scaledsampledf, centers = k, nstart = 25)

#Add the K-Means cluster labels
scaledsampledf$kmeans_cluster <- as.factor(km_result$cluster)

#Visualise clusters
ggplot(scaledsampledf, aes(x = floor_area_sqm, y = resale_price, color = kmeans_cluster)) +
  geom_point(alpha = 0.5) +
  labs(title = "Cluster Visualization (1990-1991 Flats)", x = "Floor Area (scaled)", y = "Resale Price (scaled)") +
  theme_minimal()
```
Find out value of WCSS at k=3
```{r}
km_result$tot.withinss
```
Using Hierarchical Clustering
```{r}
#Compute distance matrix
dist_matrix <- dist(scaledsampledf, method = "euclidean")
#Perform hierarchical clustering
hc_result <- hclust(dist_matrix, method = "ward.D2")

#Plot dendrogram
plot(hc_result, labels = FALSE, main = "Hierarchical Clustering Dendrogram (1990-1991 Flats)", sub = "", xlab = "")
#Add rectangles to indicate clusters
rect.hclust(hc_result, k = k, border = "red")

#cut the tree according to k=3
k <- 3
#Add cluster label to scaledsampledf
scaledsampledf$hierarchical_cluster <- cutree(hc_result, k)  

#Check scaledsampledf
head(scaledsampledf) 
```

Comparison for k-means and hierarchical clusters
```{r}
kmeans_cluster_sizes <- table(km_result$cluster)
hc_cluster_sizes <- table(scaledsampledf$hierarchical_cluster)

# Print cluster sizes for comparison
cat("K-Means Cluster Sizes:\n")
print(kmeans_cluster_sizes)
cat("\nHierarchical Cluster Sizes:\n")
print(hc_cluster_sizes)
```
Find the means of each variable in each clustering technique
```{r}
#k-means 
aggregate(scaledsampledf[, c("flat_type_numeric", "floor_area_sqm", "resale_price")],
          by = list(kmeans_cluster = scaledsampledf$kmeans_cluster), 
          FUN = mean)

#hierarchical 
aggregated_data <- aggregate(scaledsampledf[, c("flat_type_numeric", "floor_area_sqm", "resale_price")],
                             by = list(hierarchical_cluster = scaledsampledf$hierarchical_cluster), 
                             FUN = mean)
#Display aggregated data for hierarchical clustering
print(aggregated_data)
```

Reassign hierarchical clusters based on results
```{r}
#Create a mapping to swap Hierarchical cluster labels 2 and 3
scaledsampledf$cluster_swapped <- scaledsampledf$hierarchical_cluster

#Swap the cluster 2 to 3 and cluster 3 to 2
scaledsampledf$cluster_swapped[scaledsampledf$hierarchical_cluster == 2] <- 3
scaledsampledf$cluster_swapped[scaledsampledf$hierarchical_cluster == 3] <- 2

#Ensure 'cluster_swapped' is a factor
scaledsampledf$cluster_swapped <- as.factor(scaledsampledf$cluster_swapped)

#Check if it is reassigned correctly
table(scaledsampledf$cluster_swapped)
```
Visualise the updated hierarchical clustering
```{r}
fviz_cluster(list(data = scaledsampledf[, c("flat_type_numeric", "floor_area_sqm", "resale_price")], 
                  cluster = scaledsampledf$cluster_swapped),
             geom = "point", stand = FALSE, ellipse = TRUE,
             main = "Cluster Visualization (Hierarchical Clustering with Swapped Labels)")
```
Compare sizes of reassigned clusters
```{r}
#Hierarchical clusters
newhc_cluster_sizes <- table(scaledsampledf$cluster_swapped)
cat("\nReassigned Hierarchical Cluster Sizes:\n")
print(newhc_cluster_sizes)
#k-means clusters
cat("\nK-Means Cluster Sizes:\n")
print(kmeans_cluster_sizes)
```
Silhuoette Scores; higher values indicates better-defined clusters
```{r}
#k-means 
kmeans_silhouette <- silhouette(km_result$cluster, dist_matrix)
#Extract the silhouette scores and calculate the mean
kmeans_silhouette_score <- mean(kmeans_silhouette[, 3])
#Print the mean silhouette score
print(kmeans_silhouette_score)

#hierarchical 
hc_silhouette <- silhouette(as.numeric(scaledsampledf$cluster_swapped), dist_matrix)
#Extract the silhouette scores and calculate the mean
hc_silhouette_score <- mean(hc_silhouette[, 3])
#Print the mean silhouette score
print(hc_silhouette_score)
```
Perform PCA on scaled data
```{r}
#Perform PCA 
numeric_data <- scaledsampledf[, c("flat_type_numeric", "floor_area_sqm", "resale_price")]
pca_result <- prcomp(numeric_data, scale. = TRUE)
#Convert PCA result to data frame
pca_data <- as.data.frame(pca_result$x)
#Visualise the clusters using PCA (K-Means clusters)
fviz_cluster(km_result, data = pca_data,
             geom = "point", stand = FALSE, ellipse = TRUE,
             main = "K-Means Clustering Visualization (PCA Reduced)")
```
Plot boxplot for Floor Area (sqm) by Flat Type
```{r}
ggplot(df_filtered, aes(x = factor(flat_type), y = floor_area_sqm, fill = factor(flat_type))) +
  geom_boxplot() +
  labs(title = "Floor Area Distribution by Flat Type", 
       x = "Flat Type", 
       y = "Floor Area (Square Meters)") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")
```
Plot boxplot for Resale Price vs Flat Type
```{r}
ggplot(df_filtered, aes(x = factor(flat_type), y = resale_price, fill = factor(flat_type))) +
  geom_boxplot() +
  labs(title = "Resale Price (SGD) by Flat Type", 
       x = "Flat Type", 
       y = "Resale Price (SGD)") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")
```